
% ----------------------------------------------------------------
% AMS-LaTeX Paper ************************************************
% **** -----------------------------------------------------------
\documentclass[11pt]{amsart}
\usepackage{graphicx, amssymb,amsfonts,amsmath,amsthm,newlfont,mathtools} % mathabx
\usepackage{epsfig,url}
\usepackage[usenames,dvipsnames]{color}
\usepackage{enumerate}
\usepackage[colorlinks=true,linkcolor=red,citecolor=blue]{hyperref}
\usepackage{color}
\usepackage{stmaryrd}         % Crochet double barre (entiers)
\usepackage{xargs}
\usepackage{overpic}

\input{aux/commands}
\graphicspath{{figures/}} % les repertoires ou se trouvent les figures

% -----------------------------------------------------------------

\title{The combinatorics of the permutahedron diagonals}

\author[B. Delcroix-Oger]{B\'er\'enice Delcroix-Oger}
\address{Institut Montpelli\'erain Alexander Grothendieck, Universit\'e de Montpellier, France}
\email{berenice.delcroix-oger@umontpellier.fr}

\author[M. Josuat-Verg\`es]{Matthieu Josuat-Verg\`es}
\address{}
\email{}

\author[G. Laplante-Anfossi]{Guillaume Laplante-Anfossi}
\address{School of Mathematics and Statistics, The University of Melbourne, Victoria, Australia}
\email{guillaume.laplanteanfossi@unimelb.edu.au}

\author{Vincent Pilaud}
\address[Vincent Pilaud]{CNRS \& LIX, \'Ecole Polytechnique, Palaiseau}
\email{vincent.pilaud@lix.polytechnique.fr}
\urladdr{\url{http://www.lix.polytechnique.fr/~pilaud/}}

\author[K. Stoeckl]{Kurt Stoeckl}
\address{School of Mathematics and Statistics, The University of Melbourne, Victoria, Australia}
\email{kstoeckl@student.unimelb.edu.au}

\date{\today}

\subjclass[2010]{Primary [...]; Secondary 18M70} 

\keywords{Polytopes [...]}


%\thanks{G. L.-A. was supported by }

%========================================
\begin{document}
%========================================

\begin{abstract}
The purpose of this article is to provide a systematic combinatorial-geometric study of cellular diagonals on the permutahedra. 
First, we give general enumeration results for the faces of any geometric diagonal, using a theorem of Zavlasky on the M\"obius function of hyperplane arrangements.
The numbers obtained involve covering trees of bipartite graphs and Catalan families. 
Second, we prove that there are only two operadic families of diagonals on the permutahedra, and we give a bijection of their facets with planar bipartite trees and characterize their vertices as pattern-avoiding pairs of permutations.
A corollary of our study is that the Saneblidze--Umble diagonal can be recovered by a choice of vectors in the fundamental hyperplane arrangements of the permutahedra, resolving a conjecture of the third-named author.  
\end{abstract}


\maketitle


\setcounter{tocdepth}{2}
\tableofcontents

\input{sec/introduction}
\input{sec/fvectors}
\input{sec/preliminaries}
\input{sec/bijections}
\input{sec/tables}




\bigskip 

\emph{Acknowledgements.}    
Hugh Thomas (everything started in Montr\'eal), Sylvie Corteel for hyperplane arrangements

We would like to thank the Max Planck Institute for Mathematics in Bonn where part of this work was carried out. 
The third author was supported by the Australian Research Council Future Fellowship FT210100256.

\bibliographystyle{amsalpha}

\bibliography{Poissons}

\end{document}